### 2024-08-26 ms.qmd
Updates

- Quantitative abilities and its importance
The knowledge of quantities and the relation among them is one of the highest properties of cognition (Reznikova 2018). Among those numerical competences, the simplest ability is to make judgments and comparisons between “more” versus “less” (Reznikova 2007; Stancher et al. 2015), which is widespread in the animal kingdom (Vonk and Beran 2012; Beran et al. 2014). Sensitivity to relative differences in quantity might be adaptive in several situations (Nieder 2017, 2018). For instance, female lions (Panthera leo) seem to use numerical assessment of possible rivals to cope with or avoid social contests (McComb et al. 1994). Moreover, the ability to correctly estimate quantities is also associated with, for example, navigation (Chitka and Geiger 1995), choosing larger groups of potential mates (Carazo et al. 2009) and/or social companions (Bisazza et al. 2010), or increasing foraging efficiency (Uller et al. 2003).
In this sense, the literature is plenty of studies employing the spontaneous choice approach in many animals (e.g., see Carazo et al. 2009; Pahl et al. 2013 for invertebrates; Bisazza et al. 2010; Gómez-Laplaza and Gerlai 2013; Mehlis et al. 2015 for fishes; Uller et al. 2003; Krusche et al. 2010; Stancher et al. 2015 for amphibians; Garland et al. 2012 for birds; Ferkin et al. 2005; Abramson et al. 2011; Mahamane et al. 2014; Cox and Montrose 2016; Plotnik et al. 2019 for mammals) and also of research that explores numerical abilities with training procedures (e.g., see Gross et al. 2009; Pahl et al. 2013; Howard et al. 2018, 2019, for invertebrates; Bisazza et al. 2014 for fishes; Nieder 2017 for birds; Kilian et al. 2003; Addessi et al. 2008; Vonk and Beran 2012; Vonk 2014; Beran and Parrish 2016; Nieder 2017 for mammals).

- Different ways to test numerical abilities
There are two main approaches to study numerical abilities: spontaneous choice tests and training procedures. In spontaneous choice tests, experimenters present to subjects two groups of biologically relevant stimuli that differ in numerosity, under the assumption that animals should select the most appropriate quantity (e.g. the larger amount of food) (Agrillo and Bisazza 2014; Cox and Montrose 2016). In training
procedures, animals must learn a numerical rule to receive a reward (Agrillo and Bisazza 2014; Miletto-Petrazzini et al. 2018). Both experimental approaches have some advantages and disadvantages (reviewed in Agrillo and Bisazza 2014). In spontaneous choice tests, animals exhibit their natural behavior, enabling interspecific comparisons and the acquirement of ecological issues related to numerousness (Krusche et al. 2010; Abramson et al. 2011). However, motivation plays a key-role in spontaneous choice tests (Agrillo and Bisazza 2014), which implies that the absence of results might not be due to the lack of skills to solve the numerical task. Also, individual preferences may arise in this kind of procedures and, hence, a large sample of subjects is needed to consider inter-individual variability. Finally, it could be difficult, but necessary, to control for continuous variables such as area, density or movement (Agrillo and Bisazza 2014; Mehlis et al. 2015). Regardless, both methods are considered complementary and necessary to draw a firm picture of the numerical abilities of a species (Miletto-
Petrazzini et al. 2018).

- The role of early environment on the development of cogition
The conditions in which organisms develop can be altered by the circumstances faced by their parents. For instance, exposing threespine sticklebacks females to a model predator during ovulation, affected their offspring’s tedency to use social cues (Feng et al., 2015). Under stressful situations, animal’s sensory systems detect and transmit environmental signals to instigate adaptive physiological and behavioural adjustments – responses that are mediated by glucocorticoids (GCs) (Sapolsky et al., 2000). Glucocorticoids can be transmitted to offspring
and influence their phenotype and development t 46 hrough transgenerational effects (Crino et al., 2023). In addition to the environments experienced by parents, offspring also experience potentially stressful environmental conditions that can interact with or amplify parental effects. For example, human disturbance alters snapping turtles’ nest-site choice, ultimately impacting the temperature at which eggs are incubated (Kolbe & Janzen, 2002). Because animals’ responses to abrupt changes in temperature are mediated by GCs (Crino et al., 2023), and this response can affect animals’ decision-making while nesting, GCs transmission and early thermal environment are predicted to interact to shape offspring phenotype. Both GCs and temperature are known to exert potent effects on brain development (Amiel et al., 2017; Coomber et al., 1997; Jonson et al., 1976; Zhu et al., 2004) and cognition (Clark et al., 2014; Lui et al., 2017).
    - The role of stress hormones on the development of numerical abilities
      NOTHING DONE?
    - The role of temperature on the development of numerical abilities
      Port Jackson shark eggs were incubated and reared at current and projected end-of-century temperatures (+ 3 °C). We tested juvenile’s learning ability with a quantity discrimination task. The mortality rate of sharks reared in warm water was 41.7% compared with no mortality in the present-day sharks. Contrary to expectations, our results suggest that surviving hatchlings from the elevated-temperature group took fewer days to reach learning criterion and had a higher proportion of correct choice compared with hatchlings reared under present conditions. Additionally, this is the first data suggesting that sharks can discriminate different quantities (Vila-Pouca et al., 2019).

- Numerical abilities in reptiles
    - Learned num abilities
    - Spontaneous discrimination
In contrast, in reptiles, research of quantity discrimination is scarce and contradictory. Soldati et al. (2017) taught red-footed tortoises (Chelonoidis carbonaria) to associate cards of two colors to different volumes of reward and tested their choices afterward. In that succeeding experiment, tortoises showed a clear preference for the color card associated with a greater amount of food, thus concluding that discrimination between quantities is present in this species. Similarly, Hermann’s tortoises (Testudo hermanni) also proved to discriminate food by numerosity or size in spontaneous choice tests rating from 0.25 to 0.75 (Gazzola et al. 2018). In contrast, other studies suggest different results; in the Italian wall-lizard (Podarcis siculus), spontaneous choice tests showed no numerical discrimination abilities by this lizard (Miletto-Petrazzini et al. 2017), although learning procedures demonstrated an ability to discern between small close numbers (2 versus 4) (Miletto-Petrazzini et al. 2018). **Check new papers: Lin in turtles, Szabo in lizards, etc.**

- The role of early environment on reptiles cognition
Both GCs and temperature are known to exert potent effects on brain development (Amiel et al., 2017; Coomber et al., 1997) and cognition (Clark et al., 2014), but little is known about how they affect numerical discrimination. Most studies have reported better performance in different associative learning and motor tasks when incubated at higher temperatures (Amiel et al., 2014; Amiel & Shine, 2012; Clark et al., 2014). Studies that have reported cold-incubated lizards to learning faster (Abayarathna & Webb, 2020; Dayananda & Webb, 2017) used incubation temperatures far beyond the natural range in their hot treatments and it is unclear how such conditions relate to wild environments. Furthermore, sensorial capacities: **Studies from Burger** | **CORT and cognition in reptiles** | **CORT and brain in reptiles**

- Hypotheses
    - We expect that lizards incubated at high temperatures will have a better performance in the numerical discrimination task than those incubated at low temperatures.
    - We expect that lizards incubated with CORT will have a worse performance in the numerical discrimination task than those incubated with the vehicle.
    - We expect that the effect of CORT on the performance of the numerical discrimination task will be more pronounced in lizards incubated at low temperatures than in those incubated at high temperatures.
- Predictions:
    - Lizards incubated at high temperatures will discriminate better quantities even with low ratios (lower latencies, choose large amount more often, more interested in larger amounts - i.e. "fail" less tests).
    - Lizards incubated with CORT will discriminate worse quantities even with high ratios (higher latencies, choices not different from chance, equally interested in both amounts of food - i.e. "fail" more tests).
    - The effect of CORT on the performance of the numerical discrimination task will be more pronounced in lizards incubated at low temperatures than in those incubated at high temperatures (i.e. CORT-Cold will "fail" more tests than anyone else).

@@

The numerical discrimination test consisted in a series of two-choice trials where lizards where presented with different number of food items. Each choice was placed in one of two sides of a white 3D printed platform (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food. We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4. We never used more than four items because higher numbers are considered to be encoded by different regions of the brain [**REFs**]. The food consisted in frozen small crickets (*A. domestica*) that were thawed and dusted with calcium and multivitamin. We decided to use frozen crickets to avoid movement to affect lizards' behaviour. Lizards were fed frozen crickets during one month before the beginning of the experiments for habituation to frozen prey. Two weeks before the experiments, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B). The side where the food was placed was changed every day to avoid any side bias during the trials. However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals. We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day. The crickets used during the experiment were all the same size, but to control for other quantitative traits 
[**REFs In this regard, we must remark that, in our design, we did not control for other quantity-related continuous cues such as area, volume, density, weight, movement, or contour length that could be used for quantitative judgments (Uller et al. 2003; Krusche et al. 2010; Abramson et al. 2011; Garland et al. 2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al. 2015), neither for olfactory cues (Plotnik et al. 2019). These continuous variables usually covariate with quantity, but discrimination among them does not necessarily imply processment of numerical information (Agrillo and Bissaza 2014).**]
, we changed the orientation of the crickets as shown in @fig-Methods D. In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).

@@

#### Statistical analyses  
INT This gave as a gradient from -300 to +300 where the minimum represents interest in the lower amount of food and the maximum shows a preference towards higher number of crickets, while 0 shows no discrimination between them.
**We performed the analyses for each species separately. We fitted two different Bayesian multilevel models using the brm function from the _brms_ package [@burkner2017brms] using an R extension (version 2.8.2) [@R]. We ran four parallel MCMC chains of 3000 iterations for each model, with a warmup period of 1000 iterations. We modelled correct choices [correct (1) or not (0)] as the response variable, and trial, hormone (CORT versus Control), incubation temperature (Cold versus Hot), and their interaction as fixed factors. We included a random intercept and slope (trial) for each lizard in our models. We also incorporated the clutch identity as a random factor. We used the posterior distributions of parameters from these models to test for differences in learning rate between treatments and species. Learning slopes were obtained using the ‘trial’ estimates and its interaction with hormone and temperature treatments. Slope estimates greater than zero were considered as evidence of learning, while those less or equal to zero were not. We calculated the probability that slopes or contrasts between slopes differed from zero (p~mcmc~) using the posterior distributions. We considered statistical significance if p~mcmc~  < 0.05.**

#### Ethical note
Subjects were collected from a breeding colony kept in the lab since 2019. All animals used in this experiment were provided humane laboratory housing, with thermorregulation opportunity, light, and humidity conditions (see above for details). Lizards in the colony received five-seven crickets per individual three times a week while the subjects in the experiment were fed once per day. All lizards were handled minimally to avoid stress. Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33). Upon completion of experimental trials, lizards were kept in their enclosure until they were euthanized following standard humane procedures.

@@

```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
source(here("R", "func.R"))
#
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
side <- c("Right", "Left")
#
n_list <- list()
#
for(k in 1:length(hormone)){
  for(l in 1:length(temperature)){
    for(i in 1:length(side)){
      n_list[[paste(hormone[k], temperature[l], side[i], sep = "_")]] <- sample(hormone[k], temperature[l], side [i])
    }
  }
}
```

```{r, models}
#| label: models
# Fitting the model and extraction of posteriors 
refit = FALSE
if(refit){
  # Get the formula and family of variables
  lat <- bf(log(latency) ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = gaussian())
  choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
  int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
    family = gaussian())
  # Fit the model
  model <- brm(lat + choice + int,
                chains = 4, cores = 4, iter = 3000, warmup = 1000, control = list(adapt_delta = 0.99, max_treedepth =12),
                data = data_num)
    # Write the model to a file
    saveRDS(model, file = here("output/models/model_num.rds"))
  } else {
      # Read the model from a file
      model <- readRDS(file = here("output/models/model_num.rds"))
  } 
# 
```

@@

Include here general text: Results are in Table 1 and figure X.

Preliminary results showed: talk about sex and age, refer to summary (prepare that), and explain how did we control by that.

@@

Results: Latency, Choice, and Interest for each test and treatment, references to fig x and Table 1. Basically, no effects of treatment on the performance of the numerical discrimination task. no discrimination either (refer to Table in Suppl. Material).

@@

## Discussion
- Animals do not discriminate independently of the treatment or the ratio of the numerical task.
  - Other studies about discrimination
  - The importance of context
    - Learned VS spontaneous discrimination
There are two main approaches to study numerical abilities: spontaneous choice tests and training procedures. In spontaneous choice tests, experimenters present to subjects two groups of biologically relevant stimuli that differ in numerosity, under the assumption that animals should select the most appropriate quantity (e.g. the larger amount of food) (Agrillo and Bisazza 2014; Cox and Montrose 2016). In training
procedures, animals must learn a numerical rule to receive a reward (Agrillo and Bisazza 2014; Miletto-Petrazzini et al. 2018). Both experimental approaches have some advantages and disadvantages (reviewed in Agrillo and Bisazza 2014). In spontaneous choice tests, animals exhibit their natural behavior, enabling interspecific comparisons and the acquirement of ecological issues related to numerousness (Krusche et al. 2010; Abramson et al. 2011). However, motivation plays a key-role in spontaneous choice tests (Agrillo and Bisazza 2014), which implies that the absence of results might not be due to the lack of skills to solve the numerical task. Also, individual preferences may arise in this kind of procedures and, hence, a large sample of subjects is needed to consider inter-individual variability. Finally, it could be difficult, but necessary, to control for continuous variables such as area, density or movement (Agrillo and Bisazza 2014; Mehlis et al. 2015). Regardless, both methods are considered complementary and necessary to draw a firm picture of the numerical abilities of a species (Miletto-
Petrazzini et al. 2018). the Italian wall-lizard (Podarcis siculus), spontaneous choice tests showed no numerical discrimination abilities by this lizard (Miletto-Petrazzini et al. 2017), although learning procedures demonstrated an ability to discern between small close numbers (2 versus 4) (Miletto-Petrazzini et al. 2018)
    - Ecological context: food VS predators/conspecifics/etc 
Something with fishes, food, and conspecifics (**check our paper**)

- Treatment did not affect the discrimination capacity or the decision making
  - Other studies showing the effect of temperature on discrimination
  - Other studies about the effects of hormones on discrimination
  - Studies showing the effect of temperature and CORT on lizards' cognition
  - The 'two systems'
Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review see Cantlon et al. 2009; Hyde 2011). **Are they encoded in different regions? Brin-nuclei-dependent effects of CORT and temp**

- Conclusions

>>>>

The ability to discern between more and less quantities is a basic cognitive skill widely observed across the animal kingdom (**Stancher et al. 2015; Vonk and Beran 2012; Beran et al. 2014**), and is considered to be adaptive in different ecological contexts (**Nieder 2017, 2018**). Female lions (_Panthera leo_), for example, can use numerical information to assess the risk of confronting rival groups (**McComb et al. 1994**); but other animals can employ quantitative abilities for navigation (**Chitka and Geiger 1995**), mate selection (**Carazo et al. 2009**) or foraging (**Uller et al. 2003**). Given the pervasive advantages of numerical and quantity discrimination, research has extensively explored the ability of different species to differentiate between quantities under different contexts and paradigmas. Some studies focus on the type of cues involved in quantitative competences (**REFs**), others on the context where animals may use these abilities (**REFs above**), and others in the neural mechanisms underlying quantitative and numerical discrimination (**REFs**). The later have demonstrated that non-verbal numerical representations activate an extensive cortical network, with the prefrontal cortex and the intraparietal sulcus being particularly relevant (**Nieder, 2005**). This link between numerosity and brain physiology underscores the importance of understanding how brain development influences numerical abilities.

Animal's brain is particularly sensitive to environmental inputs during early stages of development [@coomber_independent_1997; @zhu_prenatal_2004], with long-lasting effects on cognitive abilities [@zhu_prenatal_2004; @amiel_egg_2014]. For instance, increases in glucocorticoids (GCs) - the main groups of hormones mediating the response to stressors [@sapolsky_how_2000] - during early life can alter animals' performance in spatial, associative, or reversal learning tasks [@szuran_water_1994; @zhu_prenatal_2004; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016; @lui2017chronic]. These effects have been shown to vary depending on factors such as dose, subjects' sex, or type of cognitive test employed [@szuran_water_1994 @farrell_developmental_2015-learn; @bebus_associative_2016], but little is known about how exposure to GCs during development can affect quantitative abilities. Similarly, thermal early environment plays a prominent role in brain development and cognitive abilities in ectotherms [@coomber_independent_1997;@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017; @vila_pouca_incubation_2018; @vila_pouca_quantity_2019]. For example, Port Jackson shark incubated at elevated-temperatures learned faster in a quantity discrimination task [@vila_pouca_quantity_2019]. However, the effects of early thermal environment on numerical abilities in other taxa remain unknown. In this sense, not only is it well-established that early-life GCs and thermal environment independently affect cognitive abilities, but these factors are also predicted to interact. In vertebrates, abrupt changes in temperature can lead to increased GC levels which, in turn, may elevate GCs in their offspring [@racic_effects_2020; @Crino_2023]. Simultaneously, these offspring are directly influenced by the same thermal environment, resulting in significant interactions that can impact cognitive development. As such, and given the ecological importance of quantitative abilities in nature [**Nieder 2017, 2018**], it is crucial to study the combined effects of early-life GCs and thermal environment on quantity and numerical discrimination. Here, we investigate the effects of prenatal exposure to corticosterone (CORT) - the main GC in reptiles [@Crino_2023] - and thermal environment on the ability of lizards to discriminate between different number of food items using the Common garden skink (_Lampropholis guichenoti)_ as a model species.

Traditionally, the study of numerical abilities has been mainly focused on mammals, birds, and fishes (**Agrillo and Bisazza 2014; Nieder 2017**), but recent research has shown that reptiles are also competent in numbers. Red-footed turtles (_Chelonoidis carbonaria_) rapidly learned to associate colour to higher volumes of food (**Soldati et al. 2017**), Hermann’s tortoises (_Testudo hermanni_) proved to discriminate between big and small amounts off food in a spontaneous choice test (**Gazzola et al. 2018**), and freshwater turtles (_Trachemys scripta_ and _Mauremys sinensis_) showed high skills when tested in a trained quantity discrimination task (**Lin et al. 2021, 2024**). In lizards, __Podarcis siculus_ did not discriminate between different amounts of food in a spontaneous choice test (**Miletto-Petrazzini et al. 2017**), but were successful in a trained quantity discrimination task (**Miletto-Petrazzini et al. 2018**), rock lizards (_Iberolacerta cyreni_) chose bigger quantities of crickets when presented 5 versus 10 items [@recio_prey_2021], and Gidgee skinks (_Egernia stokesii_) used numbers to select the larger amount of food in a spontaneous choice test [@szabo_spontaneous_2021]. Inded, the literature evidences basic quantitative abilities in reptiles, that may vary with the type of cue and test employed. However, the question of how early environment can affect numerical abilities remains open as, to our knowledge, there is no study investigating how early life conditions impact numerical discrimination in this group. Incubating lizards at elevated temperatures has been shown to improve cognitive abilities in consistency with elevated neuronal density in some areas of the brain [@amiel_hotter_2012; @amiel_egg_2014; @clark_colour_2014; @amiel_effects_2017]; and thermal environment impaired cognition only under incubation temperatures beyond lizards' natural range [@dayananda_incubation_2017 or @abayarathna_effects_2020]. Our understanding of the effects of prenatal CORT on cognitive abilities in reptiles is less understood, but in other taxa, high levels of GCs are tipically related to cognitive impairments [@szuran_water_1994; @zhu_prenatal_2004].

In our study, we manipulated CORT concentration in eggs and incubated them under one of two temperature regimes (Cold -- 23ºC ± 3ºC or Hot -- 28ºC ± 3ºC) in a 2x2 factorial design. After hatching, we assessed the lizards' numerical discrimination abilities in a spontaneous choice test, where they selected between different numbers of crickets. We hypothesized that lizards incubated at higher temperatures would outperform those incubated at lower temperatures in the numerical discrimination task. Additionally, we predicted that increased CORT levels would impair numerical competence. Moreover, we expected the negative impact of CORT on numerical performance to be more pronounced in lizards incubated at lower temperatures compared to those in higher temperatures. In line with Weber's law [**REFs**], we further anticipated that lizards incubated at higher temperatures would successfully discriminate between quantities even at high ratios (e.g. 3 VS 4), whereas event the lowest ratios (1 VS 4) could be challenging for those incubated with CORT.

@@

The numerical discrimination test consisted in a series of two-choice trials where lizards where presented with different number of food items. Each choice was placed in one of two sides of a white 3D printed platform (see @fig-Methods B, C) and then covered with a transparent Petri dish to prevent the lizard from obtaining the food. We performed five tests that differed in the ratio between both choices: 1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4. We never used more than four items because higher numbers are considered to be encoded by different regions of the brain [**REFs**]. The food consisted in frozen small crickets (*A. domestica*) that were thawed and dusted with calcium and multivitamin. We decided to use frozen crickets to avoid movement to affect lizards' behaviour. Lizards were fed frozen crickets during one month before the beginning of the experiments for habituation to frozen prey. Two weeks before the experiments, to habituate lizards to the platform and the Petri dish, the frozen cricket that was placed on top of the Petri dish in one of the sides of the platform, simulating the conditions of the experiment (see @fig-Methods B). The side where the food was placed was changed every day to avoid any side bias during the trials. However, to further control for that, the side where we placed the higher number of crickets was randomised between trials and individuals. We also counterbalanced the amount of individuals per treatment tested in the same type of trial each day. The crickets used during the experiment were all the same size, but to control for other quantitative traits 
(**Uller et al. 2003; Krusche et al. 2010; Abramson et al. 2011; Garland et al. 2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al. 2015**), we changed the orientation of the crickets as shown in @fig-Methods D. In addition, after the experiments, we made another set of trials to test whether the orientation of the crickets could affect lizards' choice, but we found no effect of orientation (see Supplementary Material - Prey orientation test).

@@

#### Statistical analyses  
In our analyses, we fitted a multivariate Bayesian multilevel model using the brm function from the _brms_ package [@burkner2017brms] in R extension (version 2.8.2) [@R]. We ran four parallel MCMC chains of 4000 iterations with a warmup period of 1000 iterations. We modelled the three main recorded behaviour: Latency (log-transformed), Choice, and Interest (see above) as the response variables, and test (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4), hormone (CORT versus Control), incubation temperature (Cold versus Hot) and their interaction as fixed factors. We also included in our predictors sex, and age (centered to 0). We included lizard indentity and clutch as random factors. We used the posterior distributions of parameters from these models to test for differences in the ability to discriminate between different number of crickets between treatments and species, for each of the numerical tests. We calculated the contrast between treatments for each test using p~mcmc~ values, and we also estimated whether predicted Choice and Interest differed from 0.5 or zero respectively using p~mcmc~. We considered statistical significance if p~mcmc~ < 0.05.

#### Ethical note
Subjects were collected from a breeding colony kept in the lab since 2019. All animals used in this experiment were provided humane laboratory housing, with thermorregulation opportunity, light, and humidity conditions (see above for details). Lizards in the colony received five-seven crickets per individual three times a week while the subjects in the experiment were fed once per day. All lizards were handled minimally to avoid stress. Experimental procedures and laboratory housing complied with Australian law and were approved by the Australian National University Animal Experimentation Ethics Committee (A2022_33). Upon completion of experimental trials, lizards were kept in their enclosure and will be euthanized following standard humane procedures once they reach the human-ending point.

@@

```{r, models}
#| label: models
# Fitting the model and extraction of posteriors 
refit = FALSE
if(refit){
  # Get the formula and family of variables
  lat <- bf(log(latency) ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = gaussian())
  choice <- bf(choice ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch), family = bernoulli(link = "logit"))
  int <- bf(compared_interest ~ test_type*temp*cort + sex + age + (1|lizard_id) + (1|clutch),
    family = gaussian())
  # Fit the model
  model <- brm(lat + choice + int,
                chains = 4, cores = 4, iter = 4000, warmup = 1000, control = list(adapt_delta = 0.99, max_treedepth =12),
                data = data_num)
    # Write the model to a file
    saveRDS(model, file = here("output/models/model_num.rds"))
  } else {
      # Read the model from a file
      model <- readRDS(file = here("output/models/model_num.rds"))
  } 
# 
```

@@

Our results showed an effect of sex and age on Latency (mean estimated effect of sex = `r {format_dec(mean(posteriors$b_loglatency_sexm), 2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_sexm, 0.975), 2)}`]; mean estimated effect of age = `r {format_dec(mean(posteriors$b_loglatency_age),2)}`, 95% CI = [`r {format_dec(quantile(posteriors$b_loglatency_age, 0.025), 2)}`, `r {format_dec(quantile(posteriors$b_loglatency_age, 0.975),2)}`]; see Table S1 in Supplementary Material), where males and older individials made, overall, faster decisions. However, we did not see such effects on Choice or Interest (see Table S1 in Supplementary Material). Nevertheless, when we analyse the effect of treatment on performance, for every variable recorded, we control for sex by averaging the predicted values for males and females, and we control for age by including the centered age in the model. These results are summarized in @fig-results and Table 1.

@@

Temperature, CORT, or their interaction contrasts did not yield any significant result in any of the numerical tests (Table 1) and we did not observe any clear pattern indicating relevant effects of early environment on any of the tests (@fig-results). Except for Hot-CORT lizards in 1 VS 3 tests or Cold-Control lizards in 2 VS 4 tests, the predicted probability of choosing the greater number of crickets first was not different from 0.5 (@fig-results; but also see Table S2 in Supplementary Material). Similarly, the predicted interest for the larger number of crickets was not different from 0 in any of the tests (@fig-results; but also see Table S2 in Supplementary Material).

@@

## Discussion
Our results indicate that lizards were not able to discriminate between different number of crickets in any of the tests, and that their performance did not vary across treatments. We did not see changes in Latency in any of the tests, and treatment did not affect lizards' time to make a decision either. Furthermore, in most of the cases, the estimated probability of choosing the option with more crickets first did not differ from random (0.5), with significant differences emerging only in two tests for two of the four treatments; yet no clear pattern was observed. Similar findings were obtained on the interest for the larger number of food items, that did not differ from zero in any of the tests or groups. These results imply that _L. guichenoti_ did not use numerical or other potential quantitative cues we were not controlling for to make foraging decisions.

In out experiment, we employed a spontaneous choice test to investigate _L. guichenoti_'s numerical abilities. This test is considered a reliable method to assess numeosity in animals, as it allows for the observation of natural behaviour and the acquirement of ecological issues related to numerousness and quantity discrimination (**REFs**). However, one of the main caveats of this approach is that motivation plays a key-role in animals' performance (**Agrillo and Bisazza, 2014**). Here, we discarded the trials where the lizards did not interact with any of the options, so the results observed are based exclusively in those where the lizards made a choice, and, as such, we do not consider motivation to be a factor affecting our results. Spontaneous choice tests are also based on the assumption that animals should select the most appropriate quantity, which may not be necessarily the largest one (**Agrillo and Bisazza, 2014**). Although the Optimal Foraging Theory predicts animals to prefer larger amounts of food, this may not hold true if increased handling costs are involved (**REFs**). Nonetheless, in that scenario, we would still expect animals to exhibit a preference, even if for the smaller amount of food, or to observe some influence of age on their choice or interest, as older animals are typically larger and may be more adept at handling a greater number of prey. Finally, spontaneous choice tests are constraint by the researchers' ability to control for other quantitative cues that may correlate with numerosity (**Agrillo and Bisazza, 2014; Uller et al. 2003; Krusche et al. 2010; Abramson et al. 2011; Garland et al. 2012; Vonk and Beran 2012; Gómez-Laplaza and Gerlai 2013; Mehlis et al. 2015**). Although we controlled for some quantitative information by rotating crickets in the option with less items to increase total area, we can not rule out the possibility that lizards could be using other quantitative cues. If lizards were using other cues exclusively for making their decision, we would still expect to see some type of discrimination between choices unless the cues were conflicting [**Cooper et al_Journal of Herpetology**]. Indeed, it could be that _L. guichenoti_ uses different quantitative cues at the same time in their foraging decisions, and, when conflicted, the preference towards one or the other varies between individuals, which may produce the same results we observed here. Alternatively, it could be that _L. guichenoti_ does not use numerical cues in their foraging decisions, and that they rely on other type of information to make their choices. Half of the studies showing numerosity in reptiles employed trained quantity discrimination tasks [**Soldati et al. 2017; Lin et al. 2021, 2024; Miletto-Petrazzini et al. 2018; Tomonaga et al. 2023 Animal Cognition**]. In the remaining studies, the relevant stimuli involved either vegetables [**Gazzola et al. 2018**; @szabo_spontaneous_2021] or large quantities (e.g., 5 vs. 10)[@recio_prey_2021]. In contrast, _Podarcis siculus_ failed to discriminate between different numbers of _Musca domestica_ larvae in a spontaneous choice test, but they showed a clear preference towards bigger larvae in a similar task [**Miletto-Petrazzini et al. 2017**]; and they were able to learn to discriminate between small close numbers in a training procedure [**Miletto-Petrazzini et al. 2018**]. This suggest that numerosity may not be relevant in foraging contexts for reptiles that eat live prey. In this regard, numerical abilities may have evolved under different situations. [**EXAMPLE FISHES THAT NOT FORAGE BUT YES SOCIAL**]. Further studies should explore numerosity in _L. guichenoti_ with a different methodology and accross different contexts. **For instance: things about higher numbers (our paper), use different type of relevant stimuli (females, conspecifics? Carazo et al and other), use training procedures.**

- Treatment did not affect the discrimination capacity or the decision making
  - Other studies showing the effect of temperature on discrimination
  - Other studies about the effects of hormones on discrimination
For instance, mammals’ hippocampi contain more GCs receptors (GRs) than other regions of the brain (Siegel et al., 2006), which may be the main cause of overall impairments inspatial learning in rats but not in other cognitive domains (Haq et al., 2021).
  - Studies showing the effect of temperature and CORT on lizards' cognition
  - The 'two systems'
The second non-verbal system is thought to be an object tracking system (also known as ‘subitizing’) that yields relatively discrete representations. This system keeps track of a small number of items (originally set to 6 (REF. 32), but later restricted to 3 or 4 items) by assign-ing ‘markers’ (also known as ‘fingers’,‘pointers’,‘symbols’, 'tokens’ or ‘files’) to individual elements (**Nieder, 2005 Nature Reviews**)
Researchers have suggested that numerousness can be processed by two types of systems: one system for small sets (≤ 4) that works by tracking and encoding objects as individual entities; and an analogous system that operates with large numerosities by estimation processes (for a review see Cantlon et al. 2009; Hyde 2011). **Are they encoded in different regions? Brin-nuclei-dependent effects of CORT and temp**

- Conclusions
Our results indicate that _L. guichenoti_ does not use numerical to make foraging decisions. This outcome, however, does not imply lack of numerical abilities, since other types of experimental approaches can yield different results. We also observed that lizards' performance was not different between treatments, which suggests that the early environment did not affect their numerical abilities. However, other tests and paradigmas should be used to test this further since both CORT and temperature can have different effects on different brain nuclei.
